---
title: "All_in_one"
author: "Vladimir"
date: "2024-04-23"
output: html_document
---
# 1. Data clearing ######################################################
### The premis of the code is that based on ealiers analysis we will use data from 2020 ("2019-12-01 00:12:00", tz = "UTC") onward (because stork data quality is lesser before that and we need to keep datasets lights for processing)
### stork data is called 'combined_stork_data'(it combined all stork data available) and ff data is called 'df_ff'(geographic limitations are based on where the stork travel)
# firest set up
```{r setting up environment}
#setwd("D:/Vladimir/Vladimir/All_in_one_codes")

# List of all required packages
packages <- c("tidyverse", "data.table", "dplyr", "rlang", "lubridate", "sf", "spatialrisk", "readr", "readxl", "move2", 
              "suntools", "tidyr", "ggplot2", "Rfast", "parallel")

# Install missing packages
new_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
if (length(new_packages)) install.packages(new_packages)

# Specific check for 'rlang' version 1.1.2 or higher
if (!("rlang" %in% installed.packages()[,"Package"]) || packageVersion("rlang") < "1.1.2") {
  install.packages("rlang")
}

# Load all the required packages
lapply(packages, library, character.only = TRUE)

# Confirm the version of 'rlang' to ensure it meets the requirements
#rlang_version <- packageVersion("rlang")
#print(paste("rlang version:", rlang_version))

```

### You can CTRL-F "START HERE" if you have 'FF_STK_dataset_fusion_ONLY_combined_stork_Data_post_2020_environment.RData'
### Else follow until there (long process due to heavy datasets)

# Upload ALL the data from the files
```{r}
#Can load that instead of running to code below
#load("D:/Vladimir/FF_STKDataset_fusion/fulls_stork_file_uploaded_environment.RData")

# Define the directory containing the CSV files
directory_path <- "C:/Master thesis data/Storks"

# Get a list of all CSV files in the directory
file_paths <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Create a named vector where names are generated dynamically based on file names
file_info <- setNames(
  file_paths,
  gsub(pattern = "LifeTrack_White_Stork_", replacement = "stork_data_", 
       gsub(pattern = " ", replacement = "_", 
            gsub(pattern = "\\.csv$", replacement = "", basename(file_paths))))
)

# Columns to keep
columns_to_keep <- c("event.id", "timestamp", "location.long", "location.lat", 
                     "individual.local.identifier")

# Loop through the file_info list to read, subset, and convert data types
for (data_name in names(file_info)) {
  # Read the CSV file
  temp_data <- read.csv(file_info[data_name])
  
  # Subset columns
  if (all(columns_to_keep %in% names(temp_data))) {
    temp_data <- temp_data[, columns_to_keep]
    
    # Convert event.id to numeric
    temp_data$event.id <- as.numeric(temp_data$event.id)
    
    # Assign the cleaned data to its respective variable in the global environment
    assign(data_name, temp_data, envir = .GlobalEnv)
  } else {
    cat(sprintf("The dataset %s is missing one or more of the required columns and has been skipped.\n", data_name))
  }
}

```

### loop to modify the name ### CAN SKIP if load("D:/Vladimir/FF_STKDataset_fusion/fulls_stork_file_uploaded_environment.RData") is loaded in
```{r}
# List all datasets currently in the environment
current_names <- ls(pattern = "LifeTrack_White_Stork_")

# Function to create new names based on the current ones
new_names <- gsub(pattern = "LifeTrack_White_Stork_", replacement = "stork_data_",
                  gsub(pattern = " ", replacement = "_", current_names))

# Loop through current names and assign them new names
for (i in seq_along(current_names)) {
  assign(new_names[i], get(current_names[i]))
  rm(list = current_names[i])  # Remove the old object
}

# Optionally, list the new names to confirm the change
print(ls(pattern = "^stork_data_"))

```

# Display column names and data types for each dataset starting with "stork_data_" 
```{r}
get_column_info <- function(df) {
  data.frame(Column = names(df), Type = sapply(df, class), row.names = NULL)
}

# Retrieve all variables that start with 'stork_data_' and print their column information
dataframe_names <- ls(pattern = "^stork_data_")
dataframes <- mget(dataframe_names)

for (data_name in dataframe_names) {
  cat("\nColumn Information for:", data_name, "\n")
  print(get_column_info(dataframes[[data_name]]))
}

```

# Extract identifiers (event.id, individual.local.identifier) from datasets with names starting with "stork_data_" and add dataset labels for tracking. 
### CAN SKIP if not data clearing
```{r}
# Retrieve all variables that start with 'stork_data_' and get their identifiers
dataframe_names <- ls(pattern = "^stork_data_")
dataframes <- mget(dataframe_names)

# Function to get identifiers with dataset labels
get_identifiers_with_labels <- function(id_column) {
  all_ids <- do.call(rbind, lapply(names(dataframes), function(name) {
    data <- dataframes[[name]][, id_column, drop = FALSE]
    data$dataset <- name  # Add a column indicating the source dataset
    return(data)
  }))
  return(all_ids)
}

# Get identifiers with dataset labels
all_event_ids_with_labels <- get_identifiers_with_labels("event.id")
all_individual_ids_with_labels <- get_identifiers_with_labels("individual.local.identifier")


# Function to check for cross-dataset duplicates
check_cross_dataset_duplicates <- function(ids) {
  # Aggregate by identifier, listing datasets each identifier appears in
  aggregated_ids <- aggregate(dataset ~ ., data = ids, FUN = function(x) paste(unique(x), collapse = ", "))
  # Find identifiers that appear in more than one dataset
  cross_dataset_duplicates <- aggregated_ids[grep(",", aggregated_ids$dataset), ]
  return(nrow(cross_dataset_duplicates) > 0)
}

# Run checks
has_cross_dataset_duplicated_event.id <- check_cross_dataset_duplicates(all_event_ids_with_labels)
has_cross_dataset_duplicated_individual.local.identifier <- check_cross_dataset_duplicates(all_individual_ids_with_labels)

# Print results
cat("Are there event.id duplicated across different datasets? ", has_cross_dataset_duplicated_event.id, "\n")
cat("Are there individual.local.identifier duplicated across different datasets? ", has_cross_dataset_duplicated_individual.local.identifier, "\n")


### event.id AND individual.local.identifier ARE UNIQUE
```

#NEXT STEP IS combining all datasets in one
```{r}
# Retrieve all the dataframes
dataframe_names <- ls(pattern = "^stork_data_")
dataframes <- mget(dataframe_names)

# Combine all dataframes into one
combined_stork_data <- do.call(rbind, dataframes)

# Optionally, add row names or reset them to ensure uniqueness
rownames(combined_stork_data) <- NULL

# Check the structure of the combined dataframe
str(combined_stork_data)

# Save the combined dataset to a CSV file for further use
#write.csv(combined_stork_data, "D:/Vladimir/combined_stork_data.csv", row.names = FALSE)

# Check for duplicate in event.id within combined_stork_data
any(duplicated(combined_stork_data$event.id)) 

# çlean environment from dataset which start with stork_data
# List all objects that start with 'stork_data_'
dataframes_to_remove <- ls(pattern = "^stork_data_")

# Remove the listed data frames from the environment
rm(list=dataframes_to_remove)
rm(dataframes)
rm(temp_data)

```

# START HERE
### Based from other analysis we only want to keep dates from 2020 ("2019-12-01 00:12:00", tz = "UTC") onward so we filter that out
### Filter 'combined_stork_data' to only keep data post 2020
```{r}
# to load the result of what is done below
load("D:/Vladimir/Vladimir/FF_STK_dataset_fusion_ONLY_combined_stork_Data_post_2020_environment.RData")

# not necessary if you've loaded what's above:
# to load the df which is process by the code below
#load("D:/Vladimir/FF_STK_dataset_fusion_ONLY_combined_stork_Data_environment.RData")

# Convert the timestamp to POSIXct
combined_stork_data$timestamp <- as.POSIXct(combined_stork_data$timestamp, format="%Y-%m-%d %H:%M:%S", tz="UTC")

# Filter the data to keep only records from "2019-12-01 00:12:00" UTC onwards
combined_stork_data <- combined_stork_data[combined_stork_data$timestamp >= as.POSIXct("2019-12-01 00:12:00", tz="UTC"), ]


### Filter combined_stk_data to not have NA in latitudes and longitudes and transform coordinates in sf
# Modify the combined stork tracking data for spatial analysis
combined_stork_data <- combined_stork_data %>%

  # Ensure timestamp is in POSIXct format (if not already)
  #mutate(timestamp = as.POSIXct(timestamp, format="%Y-%m-%d %H:%M:%S", tz='UTC')) %>%
 
  # Remove rows where location.long or location.lat is NA
  filter(!is.na(location.long) & !is.na(location.lat)) %>%
  
  # Convert the stork tracking data to an 'sf' object to enable spatial analysis
  # Using WGS 84 (EPSG:4326) coordinate reference system
  st_as_sf(coords = c("location.long", "location.lat"), crs = 4326, agr = "constant", remove = FALSE)

# Save the combined dataset to a CSV file for further use
write.csv(combined_stork_data, "D:/Vladimir/combined_stork_data.csv", row.names = FALSE)


```

# ##### Dataclearing 2.0
# We reduce the temporal resolution of 'combined_stork_data' to one entry per hour (first observed) in 'df_stork_hourly' to reduce processing time
```{r}
# Create a new dataframe with hourly data
df_stork_hourly <- combined_stork_data %>%
  group_by(individual.local.identifier) %>%  # Group data by each bird
  mutate(hour = floor_date(timestamp, "hour")) %>%  # Create a floor date at the hourly level
  arrange(individual.local.identifier, timestamp) %>%  # Sort by bird and timestamp
  filter(!duplicated(hour)) %>%  # Keep only the first observation per hour per bird
  ungroup() %>%  # Remove the grouping
  select(-hour)  # Optionally drop the 'hour' column if no longer needed

# View the structure of the new dataset
glimpse(df_stork_hourly)

#we need to make sure that df_stork_hourly is an sf object
st_crs(df_stork_hourly)

#output should be similar to:
#Coordinate Reference System:
  #User input: EPSG:4326 
# etc...

#to save environment
#save.image("D:/Vladimir/FF_STK_dataset_fusion_ONLY_combined_stork_Data_post_2020_environment.RData")


```

# Bird data ready 
# all in: 'df_stork_hourly'


## START OF DATACLEARING FOR FOREST FIRE
## Analysis begins in 2020 based on consistent data availability from the VIIRS S-NPP 375m dataset (coverage: 2019-12-01 to 2023-12-31 in CSV format). Prior forest fire data from before 2020 has been excluded, starting analysis post "2019-12-01 00:12:00 UTC".

## FOREST FIRE download at: https://firms.modaps.eosdis.nasa.gov/download/
## Area of Interest: -19.3,-18.4,69.6,70.7
## Date Range: 2019-12-01 to 2023-12-31
## Data Format: .csv
## Data source: suomi-viirs-c2 (VIIRS S-NPP 375m)

### There's 2 .csv uploaded because some data is more rescent than other so we have to bind them together
```{r}
# can upload what is done below by doing
load("D:/Vladimir/Vladimir/All_in_one_codes/FF_STK_rdy_for_overlap_analysis.RData")


# Define the paths to the CSV files
archive_file_path <- "D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_archive_SV-C2_461780.csv"
nrt_file_path <- "D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_nrt_SV-C2_461780.csv"

# Define a function to read, transform, and return an sf data frame
preprocess_and_convert_to_sf <- function(file_path) {
  read.csv(file_path, colClasses = c("acq_time" = "character")) %>%
    select(latitude, longitude, brightness, acq_date, acq_time, confidence) %>%
    mutate(acq_datetime = as.POSIXct(paste(acq_date, acq_time), format = "%Y-%m-%d %H%M", tz = 'UTC')) %>%
    st_as_sf(coords = c("longitude", "latitude"), crs = 4326, agr = "constant", remove = FALSE)
}

# Process the first file and assign to 'df_ff'
df_ff <- preprocess_and_convert_to_sf("D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_archive_SV-C2_461780.csv")

# Read and process the second file, then combine with 'df_ff', and remove the temporary variable
df_ff <- bind_rows(df_ff, preprocess_and_convert_to_sf("D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_nrt_SV-C2_461780.csv"))

# Now 'df_ff' contains the combined data from both files

# View the structure of the newly created 'sf' object
str(df_ff)

# Save environment with both cleared datasets
#save.image("D:/Vladimir/All_in_one_codes/FF_STK_rdy_for_overlap_analysis.RData")



```


#2. Overlap analysis ################################
## Start of overlap analysis

##Prepare Spatial Coordinates and Time Data
```{r}
#need 'df_stork_hourly' and 'df_ff' in the environment first

# Transform df_stork_hourly to stk_coords_time, directly creating a regular data frame
stk_coords_time <- df_stork_hourly %>%
  select(lon = location.long, lat = location.lat, time = timestamp) %>%
  as.data.frame()

# Transform df_ff to ff_coords_time, directly creating a regular data frame
ff_coords_time <- df_ff %>%
  select(lon = longitude, lat = latitude, time = acq_datetime) %>%
  as.data.frame()

```

##Define Spatial-Temporal Matching Function for a timeframe of 12h before and after fire event
```{r}
get_extended_overlap_neighbors <- function(x, stk_coords_time, ff_coords_time, radius, daysdiff) {
  Index <- seq_len(nrow(ff_coords_time))

  # Calculate time difference in seconds
  time_diff_sec <- as.numeric(difftime(ff_coords_time$time, stk_coords_time$time[x], units = "secs"))
  #time_diff_sec <- as.numeric(difftime(ff_coords_time[,3], stk_coords_time[x,3], units = "secs"))
  
  # Updated Temporal proximity check for -0.5 days before to 0.5 days after 
  #Temporal <- (time_diff_sec >= -0.5 * 86400) & (time_diff_sec <= 0.5 * 86400) ### THIS might be wrong and we should be using:
  Temporal <- (time_diff_sec >= -daysdiff * 86400) & (time_diff_sec <= daysdiff * 86400)

  # Spatial proximity check using Haversine formula
  Spatial <- spatialrisk:::haversine_loop_cpp(ff_coords_time[,c(1,2)], lat_center=stk_coords_time[x,2], lon_center=stk_coords_time[x,1], radius=radius)
  
  if (nrow(Spatial) == 0 || length(Temporal) == 0) {
    Res <- NA
  } else {
    Out <- Index[(Index %in% Spatial$id) & Temporal]
    Res <- if (length(Out) == 0) {
      NA
    } else {
      if (length(Out) == 1) {
        Out
      } else {
        Spatial$id[Spatial$id %in% Out][which.min(Spatial$distance_m[Spatial$id %in% Out])]
      }
    }
  }
  return(Res)
}


```

# running the spatial-temporal neighbors search, extracting neighbor data, and then analyzing and saving the results

```{r}
#To be able to run the code on mutiple session of Rstudio, I have to first save the necessary environment (df_stork_hourly and df_ff) to run it on the other session
#then I will only run "# code for a single bird" with the bird I want. I have to be careful to not overload the environment
#I'll also save the whole environment in: "D:\Vladimir\Vladimir\All_in_one_codes\Full_environment_All_in_one_single_bird_overlapanalysis.Rdata" as a save switch
# then we delete except the bare minimum we need for running "# code for a single bird"

# List of essential data frames, required packages, and variables/functions
essential_objects <- c("df_stork_hourly", "stk_coords_time", "ff_coords_time", "get_extended_overlap_neighbors", "packages")

# List all objects in the environment
all_objects <- ls()

# Remove objects that are not essential
objects_to_remove <- setdiff(all_objects, essential_objects)
rm(list = objects_to_remove)

# Save the data frames, transformed data frames, and the function to a single .RData file
#save(df_stork_hourly, df_ff, stk_coords_time, ff_coords_time, get_extended_overlap_neighbors, file = "D:/Vladimir/Vladimir/All_in_one_codes/Cleaned_environment.RData")

# To load the environment if needed
#load("D:/Vladimir/Vladimir/All_in_one_codes/Cleaned_environment.RData")



```

## We need a list of birds to analyse
```{r}
# Extract unique bird identifiers
unique_birds <- unique(df_stork_hourly$individual.local.identifier)

# Count occurrences of each bird identifier
bird_counts <- table(df_stork_hourly$individual.local.identifier)

# Print the counts
print(bird_counts)


#we have to choose a bird from the 'unique_birds' list
#head(unique_birds)

#already done:
# "Adebar / DER AX142 (eobs 4360)" 
# "Alain (CK16337 FRUI, 180819)" 
# "Albert + A6V24 (eobs 3813)"
# "Alwin (E0341, e-obs 8636)"
# "Angela / DER AY470 (eobs 4001)"
# "Annika + / DER A8L98 (e-obs 3813)" --> only has 2 instances
# "Anton (eobs 3561)"
# "Bembo Sparky / DER AU891 (eobs 3042)"
# "Bodom (eobs 3608)"
# "Rosalie / DER A2M38 (e-obs 6993)"
# "Mogli / DER AW534 (e-obs 6386)"
# "Olli / DER A1P74 (e-obs 6382)"
# "Ruby / DER A1P75 (e-obs 6383)"
# "Blacky + / DER A1Y32 (e-obs 6380)"


# 130 unique birds
```


# loop which does the whole analysis for each unique 'indivudual.local.identifier' (bird)
```{r}
# Extract unique bird identifiers
unique_birds <- unique(df_stork_hourly$individual.local.identifier)

# Create a directory for output if it doesn't already exist
output_dir <- "D:/Vladimir/csv_output_All_in_one"

if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

# Process data bird by bird
for (bird in unique_birds) {
  # Subset data for the current bird
  current_stork_data <- df_stork_hourly[df_stork_hourly$individual.local.identifier == bird, ]
  current_stork_coords_time <- current_stork_data %>%
    select(lon = location.long, lat = location.lat, time = timestamp) %>%
    as.data_frame()
  
  # Find neighbors
  neighbors_100km <- sapply(seq_len(nrow(current_stork_coords_time)), 
                             FUN=function(x) get_extended_overlap_neighbors(x, current_stork_coords_time, ff_coords_time, 100000, 0.5))

  # Extract neighbor data
  neighbors_100km_data <- current_stork_data[!is.na(neighbors_100km),]
  neighbors_100km_data$Dist2Fire <- st_distance(current_stork_data[!is.na(neighbors_100km),], df_ff[neighbors_100km[!is.na(neighbors_100km)],], by_element = TRUE)

  # Ensure df_ff has an identifiable, non-spatial ID column
  df_ff <- df_ff %>% 
    mutate(fire_id = row_number()) 

  df_ff_for_join <- df_ff %>% 
    st_drop_geometry() %>%
    select(fire_id, fire_longitude = longitude, fire_latitude = latitude, fire_time = acq_datetime, fire_confidence = confidence)

  # Create the final dataset with all required information
  Final_tracks_extended <- neighbors_100km_data %>%
    mutate(Nearest_fire_ID = neighbors_100km[!is.na(neighbors_100km)]) %>%
    left_join(df_ff_for_join, by = c("Nearest_fire_ID" = "fire_id")) %>%
    mutate(TimeDiff2Fire = difftime(timestamp, df_ff_for_join$fire_time[Nearest_fire_ID]))
  
  #dropping the spatial data to wirte csv
  non_spatial_data <- st_drop_geometry(Final_tracks_extended)
  
  # Save the final tracks data to a file named after the bird
  file_name <- paste0(output_dir, "/Hourly_overlap_", gsub("[^A-Za-z0-9_\\.]", "_", bird), ".csv")
  write.csv(non_spatial_data, file_name, row.names = FALSE)

}

```

# Function to easily upload any csv
```{r}
# Function to load and transform CSV data
load_and_transform_data <- function(file_path) {
  # Read the CSV file
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert date-time fields
  data$timestamp <- as.POSIXct(data$timestamp, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  data$fire_time <- as.POSIXct(data$fire_time, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  
  # Convert Dist2Fire to numeric (assuming data is in meters)
  data$Dist2Fire <- as.numeric(data$Dist2Fire)
  
  # Convert TimeDiff2Fire to difftime in seconds
  data$TimeDiff2Fire <- as.difftime(data$TimeDiff2Fire, units = "secs")
  
  # Return the transformed data
  return(data)
}

#csv upload 
Adebar_hourly_data <- load_and_transform_data("D:\\Vladimir\\csv_output_All_in_one\\Hourly_overlap_Adebar___DER_AX142__eobs_4360_.csv")
Alain_hourly_data <- load_and_transform_data("D:\\Vladimir\\csv_output_All_in_one\\Hourly_overlap_Alain__CK16337_FRUI__180819_.csv")


```

# modify csv to add new logical statement columns: "overlap within 10km range" and "fire occuring within the last hour" and a numerical "Distance to fire within the last hour"
```{r}
# Assuming Alain_hourly_data is already loaded and transformed

# 1. Add a column to check if Dist2Fire is within 10km (10000 meters)
Alain_hourly_data$Within_10km <- Alain_hourly_data$Dist2Fire <= 10000

# 2. Add a column to check if a fire occurred within hour before and after the overlap (3600 seconds before and after)
Alain_hourly_data$Fire_within_last_hour <- abs(as.numeric(Alain_hourly_data$TimeDiff2Fire, units = "secs")) <= 3600

# Add a column to check if a fire occurred within the last hour and return the distance
Alain_hourly_data$Dist_to_fire_within_last_hour <- ifelse(
  abs(as.numeric(Alain_hourly_data$TimeDiff2Fire, units = "secs")) <= 3600, 
  Alain_hourly_data$Dist2Fire, 
  NA
)


# Check if there are any TRUE statements in the new columns
any_within_10km <- any(Alain_hourly_data$Within_10km)
any_fire_within_last_hour <- any(Alain_hourly_data$Fire_within_last_hour)
any_dist_to_fire_within_last_hour <- any(!is.na(Alain_hourly_data$Dist_to_fire_within_last_hour))


# Count the number of TRUE statements in the new columns
count_within_10km <- sum(Alain_hourly_data$Within_10km, na.rm = TRUE)
count_fire_within_last_hour <- sum(Alain_hourly_data$Fire_within_last_hour, na.rm = TRUE)
count_dist_to_fire_within_last_hour <- sum(!is.na(Alain_hourly_data$Dist_to_fire_within_last_hour))


# Print the results
cat("Are there any TRUE statements for Within_10km? ", any_within_10km, "\n")
cat("Number of TRUE statements for Within_10km: ", count_within_10km, "\n")
cat("Are there any TRUE statements for Fire_within_last_hour? ", any_fire_within_last_hour, "\n")
cat("Number of TRUE statements for Fire_within_last_hour: ", count_fire_within_last_hour, "\n")
cat("Are there any TRUE statements for Dist_to_fire_within_last_hour? ", any_dist_to_fire_within_last_hour, "\n")
cat("Number of TRUE statements for Dist_to_fire_within_last_hour: ", count_dist_to_fire_within_last_hour, "\n")




```
#AA single code bird with added columns, processing takes roughtly 2-3 seconds per bird_count$Freq (number of row with the bird in question) THIS IS WORKING!
```{r}
# Create a directory for output if it doesn't already exist
output_dir <- "D:/Vladimir/csv_output_All_in_one_plus"
if (!dir.exists(output_dir)) {dir.create(output_dir)}

start_time <- proc.time()

#output_dir <- "D:/Vladimir/csv_output_All_in_one"

# Specify the bird you want to process
bird <- "Ruby / DER A1P75 (e-obs 6383)"

# Subset data for the specified bird and convert to a data frame
current_stork_data <- df_stork_hourly[df_stork_hourly$`individual.local.identifier` == bird, ]
current_stork_coords_time <- current_stork_data %>%
  select(lon = location.long, lat = location.lat, time = timestamp) %>%
  data.frame()  # Convert to traditional data frame


# Find neighbors 
# This is the spacial analysis and very long to run
neighbors_100km <- sapply(seq_len(nrow(current_stork_coords_time)), 
                           FUN=function(x) get_extended_overlap_neighbors(x, current_stork_coords_time, ff_coords_time, 100000, 0.5))

#load df_ff for analysis below
load("D:/Vladimir/Vladimir/All_in_one_codes/df_ff.RData")

# Extract neighbor data
neighbors_100km_data <- current_stork_data[!is.na(neighbors_100km),]
#neighbors_100km_data$Dist2Fire <- st_distance(current_stork_data[!is.na(neighbors_100km),], df_ff[neighbors_100km[!is.na(neighbors_100km)],], by_element = TRUE) #dist2fire is in meters
#this one below is a test to streamline the process
neighbors_100km_data$Dist2Fire <- as.numeric(st_distance(current_stork_data[!is.na(neighbors_100km),], df_ff[neighbors_100km[!is.na(neighbors_100km)],], by_element = TRUE)) #dist2fire is in meters

# Ensure df_ff has an identifiable, non-spatial ID column
df_ff <- df_ff %>% 
  mutate(fire_id = row_number()) 

df_ff_for_join <- df_ff %>% 
  st_drop_geometry() %>%
  select(fire_id, fire_longitude = longitude, fire_latitude = latitude, fire_time = acq_datetime, fire_confidence = confidence)

#remove df_ff for the next run
rm(df_ff)

# Create the final dataset with all required information
Final_tracks_extended <- neighbors_100km_data %>%
  mutate(Nearest_fire_ID = neighbors_100km[!is.na(neighbors_100km)]) %>%
  left_join(df_ff_for_join, by = c("Nearest_fire_ID" = "fire_id")) %>%
  mutate(TimeDiff2Fire = difftime(timestamp, df_ff_for_join$fire_time[Nearest_fire_ID])) %>%
  mutate(
    Within_10km = Dist2Fire <= 10000,  # Add a column to check if Dist2Fire is within 10km (10000 meters)
    Fire_within_last_hour = abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600,  # Add a column to check if a fire occurred within hour before and after the overlap
    Dist_to_fire_within_last_hour = ifelse(abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600, Dist2Fire, NA)  # Add a column to check if a fire occurred within the last hour and return the distance
  )

  #dropping the spatial data to wirte csv
non_spatial_data <- st_drop_geometry(Final_tracks_extended)

file_name <- paste0(output_dir, "/Hourly_overlap_", gsub("[^A-Za-z0-9_\\.]", "_", bird), ".csv")
write.csv(non_spatial_data, file_name, row.names = FALSE)

end_time <- proc.time()
elapsed_time <- end_time - start_time
print(elapsed_time)


```

#AA load the new csv THIS IS WORKING
```{r}
# Function to load and transform CSV data
load_and_transform_data <- function(file_path) {
  # Read the CSV file
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert date-time fields
  data$timestamp <- as.POSIXct(data$timestamp, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  data$fire_time <- as.POSIXct(data$fire_time, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  
  # Convert Dist2Fire to numeric (assuming data is in meters)
  data$Dist2Fire <- as.numeric(data$Dist2Fire)
  
  # Convert TimeDiff2Fire to difftime in seconds
  data$TimeDiff2Fire <- as.difftime(data$TimeDiff2Fire, units = "secs")
  
  # Convert Within_10km and Fire_within_last_hour to logical
  data$Within_10km <- as.logical(data$Within_10km)
  data$Fire_within_last_hour <- as.logical(data$Fire_within_last_hour)
  
  # Convert Dist_to_fire_within_last_hour to numeric
  data$Dist_to_fire_within_last_hour <- as.numeric(data$Dist_to_fire_within_last_hour)
  
  # Return the transformed data
  return(data)
}

# Load the updated CSV data
updated_file_path <- "D:\\Vladimir\\csv_output_All_in_one_plus\\Hourly_overlap_Bembo_Sparky___DER_AU891__eobs_3042_.csv"
Bembo_hourly_data <- load_and_transform_data(updated_file_path)

# Inspect the structure of the modified dataframe
str(Adebar_hourly_data)



```


# To read all the csv we create a loop, to go faster we want to run 2 sessions at the same time, we want both session to process a similar amount of data, to do so we make a list of the birds we have (total amount minus what was already processed), order them my heaviest data (highest Freq) and tell the scripts the process every 2 lines, on scipt beggins line 1 and the other line 2.


```{r}
#To be able to run the code on mutiple session of Rstudio, I have to first save the necessary environment (df_stork_hourly and df_ff) to run it on the other session
#then I will only run "# code for a single bird" with the bird I want. I have to be careful to not overload the environment
#I'll also save the whole environment in: "D:\Vladimir\Vladimir\All_in_one_codes\Full_environment_All_in_one_single_bird_overlapanalysis.Rdata" as a save switch
# then we delete except the bare minimum we need for running "# code for a single bird"

# Clean up the environment to save RAM
essential_objects <- c("df_stork_hourly", "ff_coords_time", "stk_coords_time", "get_extended_overlap_neighbors", "packages", "birds_session_1", "birds_session_2")
all_objects <- ls()
objects_to_remove <- setdiff(all_objects, essential_objects)
rm(list = objects_to_remove)
gc()
# Save the data frames, transformed data frames, and the function to a single .RData file
#save.image(file = "D:/Vladimir/Vladimir/All_in_one_codes/Cleaned_environment_for_loop.RData")

# To load the environment if needed
#load("D:/Vladimir/Vladimir/All_in_one_codes/Cleaned_environment_for_loop.RData")



```

## define the list of birds to process and sort by frequency, then split the list into two parts
## not needed if previous code has been processed
```{r}
# Define the list of already processed birds
processed_birds <- c(
  "Adebar / DER AX142 (eobs 4360)", 
  "Alain (CK16337 FRUI, 180819)", 
  "Albert + A6V24 (eobs 3813)",
  "Alwin (E0341, e-obs 8636)",
  "Angela / DER AY470 (eobs 4001)",
  "Annika + / DER A8L98 (e-obs 3813)",
  "Anton (eobs 3561)",
  "Bembo Sparky / DER AU891 (eobs 3042)",
  "Bodom (eobs 3608)",
  "Rosalie / DER A2M38 (e-obs 6993)",
  "Mogli / DER AW534 (e-obs 6386)",
  "Olli / DER A1P74 (e-obs 6382)",
  "Ruby / DER A1P75 (e-obs 6383)",
  "Blacky + / DER A1Y32 (e-obs 6380)"
)

# Extract unique bird identifiers and their counts
bird_counts <- table(df_stork_hourly$individual.local.identifier)

# Convert to a data frame and filter out already processed birds
bird_counts_df <- as.data.frame(bird_counts) %>%
  filter(!Var1 %in% processed_birds) %>%
  arrange(Freq)  # Sort by frequency

# Split the birds into two lists for the two R sessions
birds_session_1 <- bird_counts_df$Var1[seq(1, nrow(bird_counts_df), by = 2)]
birds_session_2 <- bird_counts_df$Var1[seq(2, nrow(bird_counts_df), by = 2)]

```


#Session 1: Process every other bird starting from the first
```{r}
# Specify the output directory
output_dir <- "D:/Vladimir/csv_output_All_in_one_plus_session_2"
if (!dir.exists(output_dir)) {dir.create(output_dir)}

# Function to process each bird
process_bird <- function(bird) {
  start_time <- proc.time()
  
  # Subset data for the specified bird and convert to a data frame
  current_stork_data <- df_stork_hourly[df_stork_hourly$`individual.local.identifier` == bird, ]
  current_stork_coords_time <- current_stork_data %>%
    select(lon = location.long, lat = location.lat, time = timestamp) %>%
    data.frame()  # Convert to traditional data frame

  # Find neighbors (spatial analysis)
  neighbors_100km <- sapply(seq_len(nrow(current_stork_coords_time)), 
                            FUN=function(x) get_extended_overlap_neighbors(x, current_stork_coords_time, ff_coords_time, 100000, 0.5))
  
  # Load df_ff for analysis below (assuming df_ff is already loaded in the environment)
  load("D:/Vladimir/Vladimir/All_in_one_codes/df_ff.RData")
  
  # Extract neighbor data
  neighbors_100km_data <- current_stork_data[!is.na(neighbors_100km),]
  neighbors_100km_data$Dist2Fire <- as.numeric(st_distance(current_stork_data[!is.na(neighbors_100km),], df_ff[neighbors_100km[!is.na(neighbors_100km)],], by_element = TRUE)) # dist2fire is in meters

  # Ensure df_ff has an identifiable, non-spatial ID column
  df_ff <- df_ff %>% 
    mutate(fire_id = row_number()) 

  df_ff_for_join <- df_ff %>% 
    st_drop_geometry() %>%
    select(fire_id, fire_longitude = longitude, fire_latitude = latitude, fire_time = acq_datetime, fire_confidence = confidence)

  # Remove df_ff for the next run
  rm(df_ff)

  # Create the final dataset with all required information
  Final_tracks_extended <- neighbors_100km_data %>%
    mutate(Nearest_fire_ID = neighbors_100km[!is.na(neighbors_100km)]) %>%
    left_join(df_ff_for_join, by = c("Nearest_fire_ID" = "fire_id")) %>%
    mutate(TimeDiff2Fire = difftime(timestamp, df_ff_for_join$fire_time[Nearest_fire_ID])) %>%
    mutate(
      Within_10km = Dist2Fire <= 10000,  # Add a column to check if Dist2Fire is within 10km (10000 meters)
      Fire_within_last_hour = abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600,  # Add a column to check if a fire occurred within hour before and after the overlap
      Dist_to_fire_within_last_hour = ifelse(abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600, Dist2Fire, NA)  # Add a column to check if a fire occurred within the last hour and return the distance
    )

  # Drop the spatial data to write csv
  non_spatial_data <- st_drop_geometry(Final_tracks_extended)

  file_name <- paste0(output_dir, "/Hourly_overlap_", gsub("[^A-Za-z0-9_\\.]", "_", bird), ".csv")
  write.csv(non_spatial_data, file_name, row.names = FALSE)

  end_time <- proc.time()
  elapsed_time <- end_time - start_time
  print(elapsed_time)
  
  # Clean up the environment to save RAM
  essential_objects <- c("df_stork_hourly", "ff_coords_time", "stk_coords_time", "get_extended_overlap_neighbors", "packages", "birds_session_1", "birds_session_2")
  all_objects <- ls()
  objects_to_remove <- setdiff(all_objects, essential_objects)
  rm(list = objects_to_remove)
}

# Process each bird in session 1
for (bird in birds_session_1) {
  process_bird(bird)
}


```

# somehow, "Annika + / DER A8L98 (e-obs 3813)" is in neither lists but there's only 2 entries of this bird so let's not count it to simplify the process.
#for session 2 make sure: 
```{r}
#load("D:/Vladimir/Vladimir/All_in_one_codes/Cleaned_environment.RData")

#then

# List of essential data frames, required packages, and variables/functions
essential_objects <- c("df_stork_hourly", "stk_coords_time", "ff_coords_time", "get_extended_overlap_neighbors", "packages")

# List all objects in the environment
all_objects <- ls()

# Remove objects that are not essential
objects_to_remove <- setdiff(all_objects, essential_objects)
rm(list = objects_to_remove)

# then

# Define the list of already processed birds
processed_birds <- c(
  "Adebar / DER AX142 (eobs 4360)", 
  "Alain (CK16337 FRUI, 180819)", 
  "Albert + A6V24 (eobs 3813)",
  "Alwin (E0341, e-obs 8636)",
  "Angela / DER AY470 (eobs 4001)",
  "Annika + / DER A8L98 (e-obs 3813)",
  "Anton (eobs 3561)",
  "Bembo Sparky / DER AU891 (eobs 3042)",
  "Bodom (eobs 3608)",
  "Rosalie / DER A2M38 (e-obs 6993)",
  "Mogli / DER AW534 (e-obs 6386)",
  "Olli / DER A1P74 (e-obs 6382)",
  "Ruby / DER A1P75 (e-obs 6383)",
  "Blacky + / DER A1Y32 (e-obs 6380)"
)

# Extract unique bird identifiers and their counts
bird_counts <- table(df_stork_hourly$individual.local.identifier)

# Convert to a data frame and filter out already processed birds
bird_counts_df <- as.data.frame(bird_counts) %>%
  filter(!Var1 %in% processed_birds) %>%
  arrange(Freq)  # Sort by frequency

# Split the birds into two lists for the two R sessions
birds_session_1 <- bird_counts_df$Var1[seq(1, nrow(bird_counts_df), by = 2)]
birds_session_2 <- bird_counts_df$Var1[seq(2, nrow(bird_counts_df), by = 2)]

```

#Now we can run session 2
```{r}
# Process each bird in session 2
for (bird in birds_session_2) {
  process_bird(bird)
}

```

#session 1 was stopped, to find where we are in the process we will take the original session 1 what is in the list by what is in "D:/Vladimir/csv_output_All_in_one_plus" (output of session 1)
```{r}


```


#NOT USFULL ANYMORE - load and modify files from D:\Vladimir\csv_output_All_in_one modify them to add new columns: within_10km, Fire_within_last_hour and Dist_to_fire_within_last_hour
```{r}
# Load necessary library
library(dplyr)

# Specify the input file path and output directory
input_file_path <- "D:\\Vladimir\\csv_output_All_in_one\\Hourly_overlap_Adebar___DER_AX142__eobs_4360_.csv"
output_dir <- "D:\\Vladimir\\csv_output_All_in_one_plus"

# Function to load, transform, and update CSV data
load_transform_and_update_data <- function(file_path, output_dir) {
  # Read the CSV file
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert date-time fields
  data$timestamp <- as.POSIXct(data$timestamp, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  data$fire_time <- as.POSIXct(data$fire_time, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  
  # Convert Dist2Fire to numeric (assuming data is in meters)
  data$Dist2Fire <- as.numeric(data$Dist2Fire)
  
  # Convert TimeDiff2Fire to difftime in seconds
  data$TimeDiff2Fire <- as.difftime(data$TimeDiff2Fire, units = "secs")
  
  # Calculate the missing values
  data <- data %>%
    mutate(
      Within_10km = Dist2Fire <= 10000,  # Add a column to check if Dist2Fire is within 10km (10000 meters)
      Fire_within_last_hour = abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600,  # Add a column to check if a fire occurred within hour before and after the overlap
      Dist_to_fire_within_last_hour = ifelse(abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600, Dist2Fire, NA)  # Add a column to check if a fire occurred within the last hour and return the distance
    )
  
  # Generate output file name
  file_name <- basename(file_path)
  output_file_path <- file.path(output_dir, file_name)
  
  # Save the updated data to a new CSV file
  write.csv(data, output_file_path, row.names = FALSE)
  
  return(output_file_path)
}

# Process the CSV file
updated_file_path <- load_transform_and_update_data(input_file_path, output_dir)

cat("Updated file saved to:", updated_file_path, "\n")



```
# NOT USFULL ANYMORE - Loop to process all the files in the folder
```{r}
# Load necessary library
library(dplyr)

# Function to load, transform, and update CSV data
load_transform_and_update_data <- function(file_path, output_dir) {
  # Read the CSV file
  data <- read.csv(file_path, stringsAsFactors = FALSE)
  
  # Convert date-time fields
  data$timestamp <- as.POSIXct(data$timestamp, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  data$fire_time <- as.POSIXct(data$fire_time, format = "%Y-%m-%d %H:%M:%OS", tz = "UTC")
  
  # Convert Dist2Fire to numeric (assuming data is in meters)
  data$Dist2Fire <- as.numeric(data$Dist2Fire)
  
  # Convert TimeDiff2Fire to difftime in seconds
  data$TimeDiff2Fire <- as.difftime(data$TimeDiff2Fire, units = "secs")
  
  # Calculate the missing values
  data <- data %>%
    mutate(
      Within_10km = Dist2Fire <= 10000,  # Add a column to check if Dist2Fire is within 10km (10000 meters)
      Fire_within_last_hour = abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600,  # Add a column to check if a fire occurred within hour before and after the overlap
      Dist_to_fire_within_last_hour = ifelse(abs(as.numeric(TimeDiff2Fire, units = "secs")) <= 3600, Dist2Fire, NA)  # Add a column to check if a fire occurred within the last hour and return the distance
    )
  
  # Generate output file name
  file_name <- basename(file_path)
  output_file_path <- file.path(output_dir, file_name)
  
  # Save the updated data to a new CSV file
  write.csv(data, output_file_path, row.names = FALSE)
  
  return(output_file_path)
}

# Specify the input and output directories
input_dir <- "D:\\Vladimir\\csv_output_All_in_one"
output_dir <- "D:\\Vladimir\\csv_output_All_in_one_plus"

# Get a list of all CSV files in the input directory
csv_files <- list.files(input_dir, pattern = "\\.csv$", full.names = TRUE)

# Process each CSV file
for (file_path in csv_files) {
  updated_file_path <- load_transform_and_update_data(file_path, output_dir)
  cat("Updated file saved to:", updated_file_path, "\n")
}

```
# NOT USEFULL ANYMORE - code for single bird (old) processing takes roughtly 2-3 seconds per bird_count$Freq (number of row with the bird in question)
```{r}
# Create a directory for output if it doesn't already exist
#output_dir <- "D:/Vladimir/csv_output_All_in_one"
#if (!dir.exists(output_dir)) {dir.create(output_dir)}

start_time <- proc.time()

output_dir <- "D:/Vladimir/csv_output_All_in_one"

# Specify the bird you want to process
bird <- "Bembo Sparky / DER AU891 (eobs 3042)"

# Subset data for the specified bird and convert to a data frame
current_stork_data <- df_stork_hourly[df_stork_hourly$`individual.local.identifier` == bird, ]
current_stork_coords_time <- current_stork_data %>%
  select(lon = location.long, lat = location.lat, time = timestamp) %>%
  data.frame()  # Convert to traditional data frame


# Find neighbors 
# This is the spacial analysis and very long to run
neighbors_100km <- sapply(seq_len(nrow(current_stork_coords_time)), 
                           FUN=function(x) get_extended_overlap_neighbors(x, current_stork_coords_time, ff_coords_time, 100000, 0.5))

#load df_ff for analysis below
load("D:/Vladimir/Vladimir/All_in_one_codes/df_ff.RData")

# Extract neighbor data
neighbors_100km_data <- current_stork_data[!is.na(neighbors_100km),]
neighbors_100km_data$Dist2Fire <- st_distance(current_stork_data[!is.na(neighbors_100km),], df_ff[neighbors_100km[!is.na(neighbors_100km)],], by_element = TRUE) #dist2fire is in meters

# Ensure df_ff has an identifiable, non-spatial ID column
df_ff <- df_ff %>% 
  mutate(fire_id = row_number()) 

df_ff_for_join <- df_ff %>% 
  st_drop_geometry() %>%
  select(fire_id, fire_longitude = longitude, fire_latitude = latitude, fire_time = acq_datetime, fire_confidence = confidence)

#remove df_ff for the next run
rm(df_ff)

  # Create the final dataset with all required information
Final_tracks_extended <- neighbors_100km_data %>%
  mutate(Nearest_fire_ID = neighbors_100km[!is.na(neighbors_100km)]) %>%
  left_join(df_ff_for_join, by = c("Nearest_fire_ID" = "fire_id")) %>%
  mutate(TimeDiff2Fire = difftime(timestamp, df_ff_for_join$fire_time[Nearest_fire_ID]))
  
  #dropping the spatial data to wirte csv
non_spatial_data <- st_drop_geometry(Final_tracks_extended)

file_name <- paste0(output_dir, "/Hourly_overlap_", gsub("[^A-Za-z0-9_\\.]", "_", bird), ".csv")
write.csv(non_spatial_data, file_name, row.names = FALSE)

end_time <- proc.time()
elapsed_time <- end_time - start_time
print(elapsed_time)


```
# NOT USFULL ANYMORE - code for a single bird (old )
```{r}
# Create a directory for output if it doesn't already exist
#output_dir <- "D:/Vladimir/csv_output_All_in_one"
#if (!dir.exists(output_dir)) {dir.create(output_dir)}

output_dir <- "D:/Vladimir/csv_output_All_in_one"

# Specify the bird you want to process
bird <- "Angela / DER AY470 (eobs 4001)"

# Subset data for the specified bird and convert to a data frame
current_stork_data <- df_stork_hourly[df_stork_hourly$`individual.local.identifier` == bird, ]
current_stork_coords_time <- current_stork_data %>%
  select(lon = location.long, lat = location.lat, time = timestamp) %>%
  data.frame()  # Convert to traditional data frame

# Find neighbors 
# This is the spacial analysis and very long to run
neighbors_100km <- sapply(seq_len(nrow(current_stork_coords_time)), 
                           FUN=function(x) get_extended_overlap_neighbors(x, current_stork_coords_time, ff_coords_time, 100000, 0.5))

# Extract neighbor data
neighbors_100km_data <- current_stork_data[!is.na(neighbors_100km),]
neighbors_100km_data$Dist2Fire <- st_distance(current_stork_data[!is.na(neighbors_100km),], df_ff[neighbors_100km[!is.na(neighbors_100km)],], by_element = TRUE)

# Ensure df_ff has an identifiable, non-spatial ID column
df_ff <- df_ff %>% 
  mutate(fire_id = row_number()) 

df_ff_for_join <- df_ff %>% 
  st_drop_geometry() %>%
  select(fire_id, fire_longitude = longitude, fire_latitude = latitude, fire_time = acq_datetime, fire_confidence = confidence)

  # Create the final dataset with all required information
Final_tracks_extended <- neighbors_100km_data %>%
  mutate(Nearest_fire_ID = neighbors_100km[!is.na(neighbors_100km)]) %>%
  left_join(df_ff_for_join, by = c("Nearest_fire_ID" = "fire_id")) %>%
  mutate(TimeDiff2Fire = difftime(timestamp, df_ff_for_join$fire_time[Nearest_fire_ID]))
  
  #dropping the spatial data to wirte csv
non_spatial_data <- st_drop_geometry(Final_tracks_extended)

file_name <- paste0(output_dir, "/Hourly_overlap_", gsub("[^A-Za-z0-9_\\.]", "_", bird), ".csv")
write.csv(non_spatial_data, file_name, row.names = FALSE)

```




