---
title: "FF_STK_dataset_fusion"
author: "Vladimir"
date: "2024-04-15"
output: html_document
---
```{r setting up environment}
#super heavz
#load("D:/Vladimir/FF_STK_dataset_fusion_15_04_2024.RData")
```

```{r setting up environment}
setwd("C:/Master thesis data/FF_STK_dataset_fusion")

# Install packages if not already installed
if (!require("tidyverse")) install.packages("tidyverse")
install.packages("data.table")
install.packages("dplyr")

install.packages("ggplot2") 
# Load necessary libraries
library(data.table)
library(readr)
library(dplyr)
library(rlang)
library(lubridate)  # For working with dates and times
library(sf)  # For spatial data operations


```

###### FIRST, we start by uploading and dataclearing the stork datasets
# This loop reads specified CSV files, subsets for certain columns, converts 'event.id' to numeric, and assigns each dataset to a uniquely named variable in the global environment

# NOT NEEDED Upload all the datafrom the files (light version)
```{r}

# Define file paths and the respective names you want for the data frames
file_info <- setNames(
  c("C:/Master thesis data/Storks/LifeTrack White Stork South Africa.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Sicily.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Greece Evros Delta.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Armenia.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Poland.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Tunisia.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Kosova.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Moscow.csv",
    "C:/Master thesis data/Storks/LifeTrack White Stork Poland ECG.csv"),
  c("stork_data_South_Africa", "stork_data_Sicily", "stork_data_Greece_Evros_Delta",
    "stork_data_Armenia", "stork_data_Poland", "stork_data_Tunisia",
    "stork_data_Kosova", "stork_data_Moscow", "stork_data_Poland_ECG")
)

# Columns to keep
columns_to_keep <- c("event.id", "timestamp", "location.long", "location.lat", 
                     "individual.local.identifier", "study.name")

# Loop through the file_info list to read, subset, and convert data types
for (data_name in names(file_info)) {
  # Read the CSV file
  temp_data <- read.csv(file_info[data_name])
  
  # Subset columns
  if (all(columns_to_keep %in% names(temp_data))) {
    temp_data <- temp_data[, columns_to_keep]
    
    # Convert event.id to numeric
    temp_data$event.id <- as.numeric(temp_data$event.id)
    
    # Assign the cleaned data to its respective variable in the global environment
    assign(data_name, temp_data, envir = .GlobalEnv)
  } else {
    cat(sprintf("The dataset %s is missing one or more of the required columns and has been skipped.\n", data_name))
  }
}


```




# Upload ALL the data from the files (heavy version)
```{r}
#Can load that instead of running to code below
load("D:/Vladimir/FFSTKDataset_fusion/fulls_stork_file_uploaded_environment.RData")

# Define the directory containing the CSV files
directory_path <- "C:/Master thesis data/Storks"

# Get a list of all CSV files in the directory
file_paths <- list.files(path = directory_path, pattern = "\\.csv$", full.names = TRUE)

# Create a named vector where names are generated dynamically based on file names
file_info <- setNames(
  file_paths,
  gsub(pattern = "LifeTrack_White_Stork_", replacement = "stork_data_", 
       gsub(pattern = " ", replacement = "_", 
            gsub(pattern = "\\.csv$", replacement = "", basename(file_paths))))
)

# Columns to keep
columns_to_keep <- c("event.id", "timestamp", "location.long", "location.lat", 
                     "individual.local.identifier")

# Loop through the file_info list to read, subset, and convert data types
for (data_name in names(file_info)) {
  # Read the CSV file
  temp_data <- read.csv(file_info[data_name])
  
  # Subset columns
  if (all(columns_to_keep %in% names(temp_data))) {
    temp_data <- temp_data[, columns_to_keep]
    
    # Convert event.id to numeric
    temp_data$event.id <- as.numeric(temp_data$event.id)
    
    # Assign the cleaned data to its respective variable in the global environment
    assign(data_name, temp_data, envir = .GlobalEnv)
  } else {
    cat(sprintf("The dataset %s is missing one or more of the required columns and has been skipped.\n", data_name))
  }
}

```

### loop to modify the name (bc there was a naming problem) ### CAN SKIP if load("D:/Vladimir/FFSTKDataset_fusion/fulls_stork_file_uploaded_environment.RData") is loaded in
```{r}
# List all datasets currently in the environment
current_names <- ls(pattern = "LifeTrack_White_Stork_")

# Function to create new names based on the current ones
new_names <- gsub(pattern = "LifeTrack_White_Stork_", replacement = "stork_data_",
                  gsub(pattern = " ", replacement = "_", current_names))

# Loop through current names and assign them new names
for (i in seq_along(current_names)) {
  assign(new_names[i], get(current_names[i]))
  rm(list = current_names[i])  # Remove the old object
}

# Optionally, list the new names to confirm the change
print(ls(pattern = "^stork_data_"))

```


# Display column names and data types for each dataset starting with "stork_data_" 
```{r}
get_column_info <- function(df) {
  data.frame(Column = names(df), Type = sapply(df, class), row.names = NULL)
}

# Retrieve all variables that start with 'stork_data_' and print their column information
dataframe_names <- ls(pattern = "^stork_data_")
dataframes <- mget(dataframe_names)

for (data_name in dataframe_names) {
  cat("\nColumn Information for:", data_name, "\n")
  print(get_column_info(dataframes[[data_name]]))
}

```


# Extract identifiers (event.id, individual.local.identifier) from datasets with names starting with "stork_data_" and add dataset labels for tracking. ### CAN SKIP if not data clearing
```{r}
# Retrieve all variables that start with 'stork_data_' and get their identifiers
dataframe_names <- ls(pattern = "^stork_data_")
dataframes <- mget(dataframe_names)

# Function to get identifiers with dataset labels
get_identifiers_with_labels <- function(id_column) {
  all_ids <- do.call(rbind, lapply(names(dataframes), function(name) {
    data <- dataframes[[name]][, id_column, drop = FALSE]
    data$dataset <- name  # Add a column indicating the source dataset
    return(data)
  }))
  return(all_ids)
}

# Get identifiers with dataset labels
all_event_ids_with_labels <- get_identifiers_with_labels("event.id")
all_individual_ids_with_labels <- get_identifiers_with_labels("individual.local.identifier")


# Function to check for cross-dataset duplicates
check_cross_dataset_duplicates <- function(ids) {
  # Aggregate by identifier, listing datasets each identifier appears in
  aggregated_ids <- aggregate(dataset ~ ., data = ids, FUN = function(x) paste(unique(x), collapse = ", "))
  # Find identifiers that appear in more than one dataset
  cross_dataset_duplicates <- aggregated_ids[grep(",", aggregated_ids$dataset), ]
  return(nrow(cross_dataset_duplicates) > 0)
}

# Run checks
has_cross_dataset_duplicated_event.id <- check_cross_dataset_duplicates(all_event_ids_with_labels)
has_cross_dataset_duplicated_individual.local.identifier <- check_cross_dataset_duplicates(all_individual_ids_with_labels)

# Print results
cat("Are there event.id duplicated across different datasets? ", has_cross_dataset_duplicated_event.id, "\n")
cat("Are there individual.local.identifier duplicated across different datasets? ", has_cross_dataset_duplicated_individual.local.identifier, "\n")


### event.id AND individual.local.identifier ARE UNIQUE
```

#NEXT STEP IS combining all datasets in one
```{r}
# Retrieve all the dataframes
dataframe_names <- ls(pattern = "^stork_data_")
dataframes <- mget(dataframe_names)

# Combine all dataframes into one
combined_stork_data <- do.call(rbind, dataframes)

# Optionally, add row names or reset them to ensure uniqueness
rownames(combined_stork_data) <- NULL

# Check the structure of the combined dataframe
str(combined_stork_data)

# Save the combined dataset to a CSV file for further use
#write.csv(combined_stork_data, "D:/Vladimir/combined_stork_data.csv", row.names = FALSE)

# Check for duplicate in event.id within combined_stork_data
any(duplicated(combined_stork_data$event.id)) 

# çlean environment from dataset which start with stork_data
# List all objects that start with 'stork_data_'
dataframes_to_remove <- ls(pattern = "^stork_data_")

# Remove the listed data frames from the environment
rm(list=dataframes_to_remove)
rm(dataframes)
rm(temp_data)

```
### Based from other analysis we only want to keep dates from 2020 ("2019-12-01 00:12:00", tz = "UTC") onward so we filter that out
### Filter 'combined_stork_data' to only keep data post 2020
```{r}
# to load what is done below
# load("D:/Vladimir/FF_STK_dataset_fusion_ONLY_combined_stork_Data_post_2020_environment.RData")

# Convert the timestamp to POSIXct
combined_stork_data$timestamp <- as.POSIXct(combined_stork_data$timestamp, format="%Y-%m-%d %H:%M:%S", tz="UTC")

# Filter the data to keep only records from "2019-12-01 00:12:00" UTC onwards
combined_stork_data <- combined_stork_data[combined_stork_data$timestamp >= as.POSIXct("2019-12-01 00:12:00", tz="UTC"), ]

# free up memory
gc()

# after filtering 25900815 row


```
# To load combined_stork_data from a csv or from environment and trying the make df lighter
```{r}
#Normal long
#load("D:/Vladimir/FF_STK_dataset_fusion_ONLY_combined_stork_Data_environment.RData")

# Drop the study.name column
#combined_stork_data <- combined_stork_data %>% select(-study.name)

#very long
#combined_stork_data <- read.csv("D:/Vladimir/combined_stork_data.csv")

#install.packages("data.table")
library(data.table)
dt <- as.data.table(combined_stork_data)

#delete combined stork data to free RAM
rm(combined_stork_data)

```

# Need to adjust the data to the right type for further analysis
```{r}
#here I use 'dt' because 'combined_stork_data' is too heavy but dt can be replaces with combined_stork_data
load("D:/Vladimir/FF_STK_dataset_fusion_dt_instead_of_df.RData")


library(data.table)
library(sf)

# Convert timestamp to POSIXct format if it's not already
dt[, timestamp := as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%S", tz = 'UTC')]

# Remove rows where location.long or location.lat is NA
dt <- dt[!is.na(location.long) & !is.na(location.lat)]

# Convert the data.table to an sf object
#dt_sf <- st_as_sf(dt, coords = c("location.long", "location.lat"), crs = 4326, agr = "constant", remove = FALSE)


# Save the combined dataset to a CSV file for further use
#write.csv(combined_stork_data, "D:/Vladimir/combined_stork_data.csv", row.names = FALSE)


```


### same as the one above but with combined_stork_data and not dt
```{r}
#here I use 'dt' because 'combined_stork_data' is too heavy but dt can be replaces with combined_stork_data
#load("D:/Vladimir/FF_STK_dataset_fusion_ONLY_combined_stork_Data_environment.RData")


library(dplyr)
library(sf)

# Modify the combined stork tracking data for spatial analysis
combined_stork_data <- combined_stork_data %>%

  # Ensure timestamp is in POSIXct format (if not already)
  #mutate(timestamp = as.POSIXct(timestamp, format="%Y-%m-%d %H:%M:%S", tz='UTC')) %>%
 
  # Remove rows where location.long or location.lat is NA
  filter(!is.na(location.long) & !is.na(location.lat)) %>%
  
  # Convert the stork tracking data to an 'sf' object to enable spatial analysis
  # Using WGS 84 (EPSG:4326) coordinate reference system
  st_as_sf(coords = c("location.long", "location.lat"), crs = 4326, agr = "constant", remove = FALSE)


# Save the combined dataset to a CSV file for further use
#write.csv(combined_stork_data, "D:/Vladimir/combined_stork_data.csv", row.names = FALSE)


```


###### DATA CLEARED FOR BIRD TRACKS JUST NEED TO SCALE UP WITH ALL DATASET THEN WE CAN PROCESS THEM IN FF_STK_overlap_analysis



### START OF DATACLEARING FOR FOREST FIRE download at: https://firms.modaps.eosdis.nasa.gov/download/ use "VIIRS S-NPP 375m: Temporal Coverage: 20 January 2012 - present" or "VIIRS NOAA-20 375m: Temporal Coverage: 1 January 2020 - present"
## HERE WE USED: "VIIRS S-NPP 375m: Temporal Coverage: 2019-12-01 - 2023-12-31" in csv format

#First we need to know from know the timeframe present in the stork data by searching for the earliest and latest 'dates' 
```{r}
library(ggplot2)

# Retry the ggplot histogram
ggplot(dt, aes(x = date)) +
  geom_histogram(binwidth = 365,  # Set bin width to one year
                 color = "black", fill = "blue") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +  # Set breaks and labels to one year
  labs(title = "Distribution of Records Over Time for all Stork data",
       x = "Date",
       y = "Number of Records") +
  theme_minimal()  # Use a minimal theme

```

# Saving (pdf) plot of the distribution of data
```{r}
# To save as png in folder
# Set the path and filename for the output plot
pdf(file = "D:\\Vladimir\\Plot\\stork_histogram.pdf", width = 8, height = 6)
hist(dt$date, breaks = "months", main = "Distribution of Records Over Time for all Stork data", xlab = "Date", ylab = "Number of Records")
dev.off()  # Make sure this command is executed

# To save as png in folder
# Set the path and filename for the output plot
pdf(file = "D:\\Vladimir\\Plot\\stork_histogram_yearly.pdf", width = 8, height = 6)

ggplot(dt, aes(x = date)) +
  geom_histogram(binwidth = 365,  # Set bin width to one year
                 color = "black", fill = "blue") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +  # Set breaks and labels to one year
  labs(title = "Distribution of Records Over Time for all Stork data",
       x = "Date",
       y = "Number of Records") +
  theme_minimal()  # Use a minimal theme

dev.off()  # Make sure this command is executed


```

### The most amount of consitent data starts in 2020 so we shall start the analysis from 2020
### We will download data for the forest fire from 2020 onward for the regions 

## FOREST FIRE download at: https://firms.modaps.eosdis.nasa.gov/download/
## Area of Interest: -19.3,-18.4,69.6,70.7
## Date Range: 2019-12-01 to 2023-12-31
## Data Format: .csv
## Data source: suomi-viirs-c2 (VIIRS S-NPP 375m)

### UPDATES CODE with ff data from 2020 onward in df_ff
### there's 2 .csv uploaded because some data is more rescent than other so we have to bind them together
```{r}
# can upload what is done below by doing
# load("D:/Vladimir/FF_STK_dataset_fusion_ff_stk_2020_onward.RData")

# Load necessary libraries
library(dplyr)
library(sf)
library(lubridate)  # For handling date-times

# Define the paths to the CSV files
archive_file_path <- "D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_archive_SV-C2_461780.csv"
nrt_file_path <- "D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_nrt_SV-C2_461780.csv"

# Load necessary libraries
library(dplyr)
library(sf)
library(lubridate)  # For handling date-times

# Define a function to read, transform, and return an sf data frame
preprocess_and_convert_to_sf <- function(file_path) {
  read.csv(file_path, colClasses = c("acq_time" = "character")) %>%
    select(latitude, longitude, brightness, acq_date, acq_time, confidence) %>%
    mutate(acq_datetime = as.POSIXct(paste(acq_date, acq_time), format = "%Y-%m-%d %H%M", tz = 'UTC')) %>%
    st_as_sf(coords = c("longitude", "latitude"), crs = 4326, agr = "constant", remove = FALSE)
}

# Process the first file and assign to 'df_ff'
df_ff <- preprocess_and_convert_to_sf("D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_archive_SV-C2_461780.csv")

# Read and process the second file, then combine with 'df_ff', and remove the temporary variable
df_ff <- bind_rows(df_ff, preprocess_and_convert_to_sf("D:/Vladimir/FF_custom_region_data/2020 onward/DL_FIRE_SV-C2_2020-2024/fire_nrt_SV-C2_461780.csv"))

# Now 'df_ff' contains the combined data from both files

# View the structure of the newly created 'sf' object
str(df_ff)

```


```{r}


```

```{r}


```


```{r}


```


```{r}


```

```{r}


```

```{r}


```

```{r}


```









```{r}


```


```{r}


```
### OLD CODE WITH DATA FROM 2013/1/1 and 2020/09/16
```{r}
# Load necessary libraries
library(dplyr)
library(sf)
library(lubridate)  # For handling date-times if not already loaded

# Read the CSV file and preprocess it
custome_Wf <- read.csv("D:/Vladimir/FF_custom_region_data/fire_archive_SV-C2_456943.csv", colClasses = c("acq_time"="character")) %>%
  
  # Filter to only keep the specified columns
  select(latitude, longitude, brightness, acq_date, acq_time, confidence) %>%
  
  # Convert the acquisition date and time into a POSIXct object for accurate temporal analysis
  mutate(acq_datetime = as.POSIXct(paste(acq_date, acq_time), format="%Y-%m-%d %H%M", tz='UTC')) %>%
  
  # Convert the dataset to an 'sf' object for spatial operations, using WGS 84 CRS
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, agr = "constant", remove = FALSE)

# View the structure of the newly created 'sf' object
str(custome_Wf)
```


# dataset Upload template
```{r}
# Reading a CSV file from the Storks subdirectory
stork_data_Uzbe <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Uzbekistan.csv")

# Read each CSV file into a separate variable using the full path
stork_data_South_Africa <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork South Africa.csv")
stork_data_Sicily <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Sicily.csv")
stork_data_Greece_Evros_Delta <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Greece Evros Delta.csv")
stork_data_Armenia <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Armenia.csv")
stork_data_Poland <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Poland.csv")
stork_data_Tunisia <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Tunisia.csv")
stork_data_Kosova <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Kosova.csv")
stork_data_Moscow <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Moscow.csv")
stork_data_Poland_ECG <- read.csv("C:/Master thesis data/Storks/LifeTrack White Stork Poland ECG.csv")

```
